Supporting Information: Thermal sensitivity across forest vertical profiles: patterns, mechanisms, and ecological implications

### 2021-09-14 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

**Vertical gradients affect tree metabolism and growth.**
Whole-tree transpiration and gross primary production (GPP) increase in tandem with increasing tree size, where size is most commonly measured as diameter at breast height (DBH), but functionally linked to metrics such as foliage biomass, crown volume, and height, all of which scale with DBH [@meinzerWaterTransportTrees2001;@kunertRevisedHydrologicalModel2017; @anderson-teixeiraSizerelatedScalingTree2015].
Specifically, increases are linked to increased leaf area and in part to an increasing probability that the crown is in the canopy [@muller-landauTestingMetabolicEcology2006], where higher light availability and associated elevation of $g_s$ and $A_A$ result in higher leaf area-specific transpiration and photosynthesis (Table `r table_leaf_metabolism`).
The net foliar photosynthate production is allocated among functions including respiration, woody growth, foliar turnover, root growth and allocation to root-associated microorganisms, reproduction, defense, and storage of non-structural carbohydrates (NSCs). 
Within a stand, woody growth rates , measured as radial stem growth or biomass increment,  generally increase with tree size [e.g., @muller-landauTestingMetabolicEcology2006, Stephenson et al. 2014]. NSC reserves also increase with tree size, such that larger trees have relatively more buffer against biological and biophysical stressors [@niinemetsResponsesForestTrees2010a].

>>>>

**Vertical gradients affect tree metabolism and growth.**
Whole-tree transpiration and gross primary production (GPP) increase in tandem with increasing tree size, where size is most commonly measured as diameter at breast height (DBH), but functionally linked to metrics such as foliage biomass, crown volume, and height, all of which scale with DBH [@meinzerWaterTransportTrees2001;@kunertRevisedHydrologicalModel2017; @anderson-teixeiraSizerelatedScalingTree2015].
Specifically, increases are linked to increased leaf area and in part to an increasing probability that the crown is in the canopy [@muller-landauTestingMetabolicEcology2006], where higher light availability and associated elevation of $g_s$ and $A_A$ result in higher leaf area-specific transpiration and photosynthesis (Table `r table_leaf_metabolism`).
The net foliar photosynthate production is allocated among functions including respiration, woody growth, foliar turnover, root growth and allocation to root-associated microorganisms, reproduction, defense, and storage of non-structural carbohydrates (NSCs). 
Within a stand, woody growth rates , measured as radial stem growth or biomass increment,  generally increase with tree size [e.g., @muller-landauTestingMetabolicEcology2006, @stephensonRateTreeCarbon2014]. NSC reserves also increase with tree size, such that larger trees have relatively more buffer against biological and biophysical stressors [@niinemetsResponsesForestTrees2010a].
